\newpage

\section{Idea}

The tactics of this controller are based on two facts:

\begin{itemize}
	\item Saucers will always consume energy, no matter what they are doing.
	\item Cannon rounds are much more effective at close range.
\end{itemize}

Since saucers will always consume energy during all regimes of flight, this report believes that it is more efficient to be in a position to fire the weapon and attempt to degrade the enemy's energy at a faster rate, rather than fly defensively without firing at all. It is also more effective to be within close range of the enemy to cause more damage with the cannon, but on the other hand, the enemy's cannon will also be just as dangerous.\\
\\
Therefore, the tactics of the controller will be as discussed below.

\subsection{Flight regime}

If the enemy is far while winning, commit to engagement and fly aggressively towards the enemy. Otherwise if the enemy is close, and winning or the score is even, fly at a slow speed to stay within close range. This will also allow the enemy to pass so that the saucer can stay behind the enemy and avoid overshooting him. During winning or even scores, the saucer will always turn towards the enemy attempting to close the distance to a much more lethal firing range.

If losing, break contact and fly away from the enemy at the appropriate speed. If close, fly as fast as possible away from the enemy. If near, fly moderately, and if far, then fly at the minimum speed to conserve energy.

\subsection{Weapon employment}

Since cannon shots lose energy the further they travel, only fire at close or near ranges to maximise efficiency of energy. When close and winning, fire at maximum power. When close and the score is even, fire at medium power. When near and winning or the score is even, fire the cannon at maximum power to ensure that the cannon round will reach the enemy. During other circumstances, hold fire to conserve energy.