---
title: "CSG1132 A2 Notes"
author: "Martin Ponce"
date: "Wednesday, September 24, 2014"
output: html_document
---

```{r, echo=FALSE, include=FALSE, cache=FALSE}
require(e1071);
require(modeest);
require(foreign);
fbDataset.raw<-read.spss("./src/Ass2-dataset.sav", to.data.frame=TRUE);

```

# Intro

## Theme 1

### Research question

Is gender related to size of a user's Facebook network?

### Thesis statement

Gender is related to the size of a user's Facebook network.

## Theme 2

### Research question

Is gender related to intensity of Facebook use?

### Thesis statement

Gender is related to the intensity of Facebook use.

## Variables

### Gender and network size

- Gender
- Facebook friends
- Close friends
- Sociability

### Gender and intensity of Facebook use

- Gender
- Facebook logins
- Facebook hours

# Working with the data

To import SPSS data:

```{r, warning=FALSE}
require(foreign);

fbDataset.raw <- read.spss(
    "./src/Ass2-dataset.sav", 
    use.value.labels=TRUE, 
    to.data.frame=TRUE
);
```

Raw data set consists of 61 observations, 10 variables.

```{r}
summary(fbDataset.raw);
```

## Data cleaning

### Removing NA values

Created new object called fbDataset.rmNA with all observations with NA values removed:

```{r}
fbDataset.rmNA <- na.omit(fbDataset.raw);
```

56 observations remain.

### Removing Nil values

Removed observations that have 0 (Nil) as FB_Login:

```{r}
fbDataset.rmNil <- subset(fbDataset.rmNA, FB_Logins > 0);
```

53 observations remain.

### Removing Outliers

#### FB Logins

Removing FB Login outliers > 50

```{r}
fbDataset.rmOut <- subset(fbDataset.rmNil, FB_Logins < 50);
```

51 observations remain.

#### FB Hours

Removing FB Hours outliers > 50

```{r}
fbDataset.rmOut <- subset(fbDataset.rmOut, Hours < 50);
```

50 observations remain.

#### Close Friends

Removing Close Friends outliers > 70

```{r}
fbDataset.rmOut <- subset(fbDataset.rmOut, Close_friends < 70);
```

48 observations remain

### Create gender category

```{r}
# create new object, copy rmOut dataset
fbDataset.genderCat <- fbDataset.rmOut;

# create new category for gender
fbDataset.genderCat$Gender_category <- cut(fbDataset.genderCat$Gender, c(0, 1, 0.5), include.lowest = TRUE, labels = c(0:1));
```

### Create factors with value labels

```{r}
# create new object, copy genderCat dataset
fbDataset.factors <- fbDataset.genderCat;

# create factors with value labels
fbDataset.factors$Gender_label <- factor(
    fbDataset.factors$Gender, 
    levels = c(0, 1),
    labels = c("Female", "Male")
);
```

### Attach final dataset

```{r}
# make new set called final, assign clean data to it
fbDataset.final <- fbDataset.factors

# attach final dataset so I can refer straight to the variable
attach(fbDataset.final)
```

### Summary post data clense

```{r}
summary(fbDataset.final);
```

# Central tendency

## Gender

```{r}
mean(Gender);
```

```{r}
median(Gender);
```

mvf() = mode
```{r}
mfv(Gender);
```

```{r}
sd(Gender);
```

```{r}
skewness(Gender);
```

```{r}
kurtosis(Gender);
```

### Histogram

```{r}
h <- hist(
    Gender, 
    breaks = 2, 
    main = "Histogram: Gender", 
    ylab = "Frequency of gender", 
    xlab = "Gender",
    col = "grey",
);

# this draws the normal distribution curve over the histogram
xfit <- seq(
    min(Gender), 
    max(Gender), 
    length = 100
);

yfit <- dnorm(
    xfit, 
    mean = mean(Gender), , 
    sd = sd(Gender)
);

yfit <- yfit * diff(h$mids[1:2]) * length(Gender);

lines(
    xfit, 
    yfit, 
    col = "darkblue"
);
```

### Q-Q plot

Identifies normality of distribution.

```{r}
qqnorm(Gender, main = "Normal Q-Q Plot: Gender");
qqline(Gender, col = "red");
```

### Kernal density

```{r}
density <- density(Gender);

plot(density,
    main = "Kernal Density: Gender",
    xlab = "Gender",
    ylab = "Density",
    col = "red"
);
```

## FB Logins

```{r}
mean(FB_Logins);
```

```{r}
median(FB_Logins);
```

mvf() = mode
```{r}
mfv(FB_Logins);
```

```{r}
sd(FB_Logins);
```

```{r}
skewness(FB_Logins);
```

```{r}
kurtosis(FB_Logins);
```

### Histogram

```{r}
h <- hist(
    FB_Logins, 
    main = "Histogram: FB Logins", 
    ylab = "Frequency of FB Logins", 
    xlab = "FB Logins",
    col = "grey"
);

xfit <- seq(
    min(FB_Logins), 
    max(FB_Logins), 
    length=40
);

yfit <- dnorm(
    xfit, 
    mean=mean(FB_Logins), , 
    sd=sd(FB_Logins)
);

yfit <- yfit * diff(h$mids[1:2]) * length(FB_Logins);

lines(
    xfit, 
    yfit, 
    col = "darkblue"
);
```

### Q-Q plot

Identifies normality of distribution.

```{r}
qqnorm(FB_Logins, main = "Normal Q-Q Plot: FB Logins");
qqline(FB_Logins, col = "red");
```

The majority of points do not fall on the expected normal distribution line. The distribution of **FB Logins** is not normal. Utilise non-parametric methods.

### Log transformation

Log transformation may provide a normal distribution. Will test again for normality after transformation.

```{r}
log.FB_Logins <- log(FB_Logins)
```

Testing for normality.

```{r}
qqnorm(log.FB_Logins, main = "Normal Q-Q Plot: FB Logins \n Log transformed");
qqline(log.FB_Logins, col = "red");
```

Again, the majority of data points do not fall on the expected normal distribution line. Non-parametric methods still apply.

### Kernal density

```{r}
density <- density(FB_Logins);

plot(
    density,
    main = "Kernal Density: FB Logins",
    xlab = "FB Logins",
    ylab = "Density",
    col = "red"
);
```

Note: Can't get legend to appear.

```{r}
library(sm);
#plot.new();

male_logins <- subset(FB_Logins, Gender == 0);
female_logins <- subset (FB_Logins, Gender == 1);

logins.f <- factor(
    fbDataset.final,
    levels = c(0, 1),
    labels = c("Female", "Male")
);

# the comparison plot

sm.density.compare(
    fbDataset.final$FB_Logins,
    fbDataset.final$Gender,
    xlab = "Male vs. Female FB Logins",
    main = "FB Login distribution by gender"
);

# legend

colfill <- c(
    2:(2 + length (levels(logins.f)))
);

legend(
    235,
    0.014,
    legend = c(
        "Female",
        "Male",
        fill = colfill
    )
);
```

## FB Hours

```{r}
mean(Hours);
```

```{r}
median(Hours);
```

mvf() = mode
```{r}
mfv(Hours);
```

```{r}
sd(Hours);
```

```{r}
skewness(Hours);
```

```{r}
kurtosis(Hours);
```

### Histogram

```{r}
h <- hist(
    Hours, 
    main = "Histogram: FB Hours", 
    ylab = "Frequency of FB Hours", 
    xlab = "FB Hours",
    col = "grey"
);

xfit <- seq(
    min(Hours), 
    max(Hours), 
    length=40
);

yfit <- dnorm(
    xfit, 
    mean = mean(Hours), , 
    sd = sd(Hours)
);

yfit <- yfit * diff(h$mids[1:2]) * length(Hours);

lines(
    xfit, 
    yfit, 
    col = "darkblue"
);
```

### Q-Q plot

Identifies normality of distribution.

```{r}
qqnorm(Hours, main = "Normal Q-Q Plot: FB Hours");
qqline(Hours, col = "red");
```

The majority of points do not fall on the expected normal distribution line. The distribution of **Hours** is not normal. Utilise non-parametric methods.

### Log transformation

Log transformation may provide a normal distribution. Will test again for normality after transformation.

```{r}
log.Hours <- log(Hours)
```

Testing for normality.

```{r}
qqnorm(log.Hours, main = "Normal Q-Q Plot: FB Hours \n Log transformed");
qqline(log.Hours, col = "red");
```

Again, the majority of data points do not fall on the expected normal distribution line. Non-parametric methods still apply.

### Kernal density

```{r}
density <- density(Hours);

plot(density,
    main = "Kernal Density: FB Hours",
    xlab = "FB Hours",
    ylab = "Density",
    col = "red"
);
```

## Facebook friends

```{r}
mean(FB_friends);
```

```{r}
median(FB_friends);
```

mvf() = mode
```{r}
mfv(FB_friends);
```

```{r}
sd(FB_friends);
```

```{r}
skewness(FB_friends);
```

```{r}
kurtosis(FB_friends);
```

### Histogram

```{r}
h <- hist(
    FB_friends, 
    main = "Histogram: FB Friends", 
    ylab = "Frequency of FB Friends", 
    xlab = "FB Friends",
    col = "grey"
);

xfit <- seq(
    min(FB_friends), 
    max(FB_friends), 
    length = 100
);

yfit <- dnorm(
    xfit, 
    mean = mean(FB_friends), , 
    sd = sd(FB_friends)
);

yfit <- yfit * diff(h$mids[1:2]) * length(FB_friends);

lines(
    xfit, 
    yfit,
    col = "darkblue"
);
```

### Q-Q plot

Identifies normality of distribution.

```{r}
qqnorm(FB_friends, main = "Normal Q-Q Plot: FB Friends");
qqline(FB_friends, col = "red");
```

The majority of points do not fall on the expected normal distribution line. The distribution of **FB Friends** is not normal. Utilise non-parametric methods.

### Log transformation

Log transformation may provide a normal distribution. Will test again for normality after transformation.

```{r}
log.FB_friends <- log(FB_friends)
```

Testing for normality.

```{r}
qqnorm(log.FB_friends, main = "Normal Q-Q Plot: FB Friends \n Log transformed");
qqline(log.FB_friends, col = "red");
```

While approaching normality, the majority of data points do not fall on the expected normal distribution line. Non-parametric methods still apply.

### Kernal density

```{r}
density <- density(FB_friends);

plot(density,
    main = "Kernal Density: FB Friends",
    xlab = "FB Friends",
    ylab = "Density",
    col = "red"
);
```

## Close friends

```{r}
mean(Close_friends)
```

```{r}
median(Close_friends)
```

mvf() = mode
```{r}
mfv(Close_friends)
```

```{r}
sd(Close_friends)
```

```{r}
skewness(Close_friends)
```

```{r}
kurtosis(Close_friends)
```

### Histogram

```{r}
h <- hist(
    Close_friends, 
    main = "Histogram: Close Friends", 
    ylab = "Frequency of Close Friends", 
    xlab = "Close Friends",
    col = "grey",
    breaks = 5,
    
    # change the limit for x-axis, ylim for y-axis
    xlim = c(0, 60),
    ylim = c(0, 20),
);

xfit <- seq(
    min(Close_friends), 
    max(Close_friends), 
    length = 100
);

yfit <- dnorm(
    xfit, 
    mean = mean(Close_friends), , 
    sd = sd(Close_friends)
);

yfit <- yfit * diff(h$mids[1:2]) * length(Close_friends);

lines(
    xfit, 
    yfit, 
    col = "darkblue"
);
```

### Q-Q plot

Identifies normality of distribution.

```{r}
qqnorm(Close_friends, main = "Normal Q-Q Plot: Close Friends");
qqline(Close_friends, col = "red");
```

The majority of points do not fall on the expected normal distribution line. The distribution of **Close friends** is not normal. Utilise non-parametric methods.

### Log transformation

Log transformation may provide a normal distribution. Will test again for normality after transformation.

```{r}
log.Close_friends <- log(Close_friends)
```

Testing for normality.

```{r}
qqnorm(log.Close_friends, , main = "Normal Q-Q Plot: FB Friends \n Log transformed");
qqline(log.Close_friends, col = "red");
```

Again, the majority of data points do not fall on the expected normal distribution line. Non-parametric methods still apply.

### Kernal density

```{r}
density <- density(Close_friends);

plot(
    density,
    main = "Kernal Density: Close Friends",
    xlab = "Close Friends",
    ylab = "Density",
    col = "red"
);
```

## Sociability

```{r}
mean(Sociability)
```

```{r}
median(Sociability)
```

mvf() = mode
```{r}
mfv(Sociability)
```

```{r}
sd(Sociability)
```

```{r}
skewness(Sociability)
```

```{r}
kurtosis(Sociability)
```

### Histogram

```{r}
sociability.bins = seq(min(Sociability), max(Sociability), 1);

h <- hist(
    Sociability, 
    main = "Histogram: Sociability", 
    ylab = "Frequency of Sociability", 
    xlab = "Sociability",
    col = "grey",
    breaks = c(1, 2, 3, 4, 5),
    ylim = c(0, 30)
);

xfit <- seq(
    min(Sociability), 
    max(Sociability), 
    length = 100
);

yfit <- dnorm(
    xfit, 
    mean = mean(Sociability), , 
    sd = sd(Sociability)
);

yfit <- yfit * diff(h$mids[1:2]) * length(Sociability);

lines(
    xfit, 
    yfit, 
    col = "darkblue"
);
```

### Q-Q plot

Identifies normality of distribution.

```{r}
qqnorm(Sociability, main = "Normal Q-Q Plot: Sociability");
qqline(Sociability, col = "red");
```

The majority of points do not fall on the expected normal distribution line. The distribution of **Sociability** is not normal. Utilise non-parametric methods.

### Log transformation

Log transformation may provide a normal distribution. Will test again for normality after transformation.

```{r}
log.Sociability <- log(Sociability)
```

Testing for normality.

```{r}
qqnorm(log.Sociability, , main = "Normal Q-Q Plot: Sociability \n Log transformed");
qqline(log.Sociability, col = "red");
```

Again, the majority of data points do not fall on the expected normal distribution line. Non-parametric methods still apply.

### Kernal density

```{r}
density <- density(Sociability);

plot(
    density,
    main = "Kernal Density: Sociability",
    xlab = "Sociability",
    ylab = "Density",
    col = "red"
);
```

# Testing

Non-parametric tests chosen because the variables are non-normally distributed.

## Tests chosen

- Mann-Whitney-Wilcoxon
    - See [r-tutor.com](http://www.r-tutor.com/elementary-statistics/non-parametric-methods/mann-whitney-wilcoxon-test)
- Spearman's rho
    - See [r-bloggers.com](http://www.r-bloggers.com/non-parametric-methods-for-the-study-of-the-correlation-spearmans-rank-correlation-coefficient-and-kendall-tau-rank-correlation-coefficient/)

## Theme 1: Gender is related to network size

### Facebook friends and gender

#### Correlation coefficient

```{r}
cor(Gender, FB_friends);

# spearman
cor(Gender, FB_friends, method = "spearman");
```

```{r}
plot(Gender, FB_friends);
```

#### Mann-Whitney-Wilcoxon test

```{r warning=FALSE}
wilcox.test(FB_friends ~ Gender, alternative = "less");
```

#### Spearman's rho test

```{r warning=FALSE}
cor.test(Gender, FB_friends, alternative = "less", method = "spearman", exact = FALSE);
```

#### Regression tests

##### Box plot

```{r}
plot(
    Gender_category, 
    FB_friends,
    main = "Box plot: \n Gender by FB Friends",
    xlab = "Gender",
    ylab = "FB Friends"
);
```

##### Scatter plot

```{r}
# plot the variables
plot(
    Gender,
    FB_friends,
    main = "Scatter plot: \n Gender by FB Friends",
    xlab = "Gender",
    ylab = "FB Friends"
);

# draw mean
abline(h = mean(FB_friends), col="darkblue");

# draw regression
abline(lm(FB_friends ~ Gender), col="red");
```

#### Summary

```{r}
summary(lm(FB_friends ~ Gender));
```

### Close friends and gender

#### Correlation coefficient

```{r}
cor(Gender, Close_friends);

# spearman
cor(Gender, Close_friends, method = "spearman");
```

#### Mann-Whitney-Wilcoxon test

```{r warning=FALSE}
wilcox.test(Close_friends ~ Gender);
```

#### Spearman's rho test

```{r warning=FALSE}
cor.test(Gender, Close_friends, alternative = "less", method = "spearman", exact = FALSE);
```

#### Regression tests

##### Box plot

```{r}
plot(
    Gender_category, 
    Close_friends,
    main = "Box plot: \n Gender by Close Friends",
    xlab = "Gender",
    ylab = "Close Friends"
);
```

##### Scatter plot

```{r}
# plot the variables
plot(
    Gender,
    Close_friends,
    main = "Scatter plot: \n Gender by Close Friends",
    xlab = "Gender",
    ylab = "Close Friends"
);

# draw mean
abline(h = mean(Close_friends), col="darkblue");

# draw regression
abline(lm(Close_friends ~ Gender), col="red");
```

#### Summary

```{r}
summary(lm(Close_friends ~ Gender));
```

### Sociability and gender

#### Correlation coefficient

```{r}
cor(Gender, Sociability);

# spearman
cor(Gender, Sociability, method = "spearman");
```

#### Mann-Whitney-Wilcoxon test

```{r warning=FALSE}
wilcox.test(Sociability ~ Gender);
```

#### Spearman's rho test

```{r warning=FALSE}
cor.test(Gender, Sociability, alternative = "less", method = "spearman", exact = FALSE);
```

#### Regression tests

##### Box plot

```{r}
plot(
    Gender_category, 
    Sociability,
    main = "Box plot: \n Gender by Sociability",
    xlab = "Gender",
    ylab = "Sociability"
);
```

##### Scatter plot

```{r}
# plot the variables
plot(
    Gender,
    Sociability,
    main = "Scatter plot: \n Gender by Sociability",
    xlab = "Gender",
    ylab = "Sociability"
);

# draw mean
abline(h = mean(Sociability), col="darkblue");

# draw regression
abline(lm(Sociability ~ Gender), col="red");
```

#### Summary

```{r}
summary(lm(Sociability ~ Gender));
```

## Theme 2: Gender is related to intensity of use 

### Gender and Facebook hours

### Sociability and gender

#### Correlation coefficient

```{r}
cor(Gender, Hours);

# spearman
cor(Gender, Hours, method = "spearman");
```

#### Correlation coefficient

```{r}
cor(Gender, Hours);
```

```{r}
plot(Gender, Hours);
```

#### Mann-Whitney-Wilcoxon test

```{r warning=FALSE}
wilcox.test(Gender, Hours, alternative = "two.sided");
```

#### Spearman's rho test

```{r warning=FALSE}
cor.test(Gender, Hours, alternative = "less", method = "spearman", exact = FALSE);
```

#### Regression tests

##### Box plot

```{r}
plot(
    Gender_category, 
    Hours,
    main = "Box plot: \n Gender by FB Hours",
    xlab = "Gender",
    ylab = "FB Hours"
);
```

##### Scatter plot

```{r}
# plot the variables
plot(
    Gender,
    Hours,
    main = "Scatter plot: \n Gender by FB Hours",
    xlab = "Gender",
    ylab = "FB Hours"
);

# draw mean
abline(h = mean(Hours), col="darkblue");

# draw regression
abline(lm(Hours ~ Gender), col="red");
```

#### Summary

```{r}
summary(lm(Gender ~ Hours));
```