\section{Research questions and thesis statements}

\subsection{Research question 1: Is gender related to a user's Facebook network size?}

Gender and its relation to SNS network size has been found to be a topic of debate. As indicated by \citet{Tifferet2014}, studies have reported conflicting results and require further investigation. The variables provided in the current dataset allows the research question and thesis statement to explore these conflicting views.

\subsection{Thesis statement 1: Gender is related to the size of a user's Facebook network size}

The number of Facebook friends, offline close friends and sociability scores will be tested against gender, seeking any correlations that may support or negate the hypothesis that gender is related to a user's Facebook network size. 

\subsection{Research Question 2: Is gender related to the amount of time a user spends on Facebook?}

The literature in the review suggests that women are attracted to the social aspect of Facebook and SNS more than men. Since the majority of activities on such sites are related to socialising, it is possible to suggest that gender is related to the amount of time a user spends on Facebook.

\subsection{Thesis statement 1: Gender is related to the amount of time a user spends on Facebook}

Reported hours will be tested against gender, seeking any correlations that may support or negate the hypothesis that gender is related to the amount of time a user spends on Facebook.

%$$s = \sqrt{\frac{1}{N-1} \sum_{i=1}^N (x_i - \overline{x})^2}$$