\section{Idea}

The main tactic for this controller is fly defensively in order to conserve energy and survive until there are a two enemies left. Turns for the most part will focus on the energy blast sensor, in order to dodge as many of them as possible. When there are only two enemy saucers left in the arena, turning will focus on the enemy saucer sensor, to track them down and shoot at them from a close distance.

However, when any power ups spawn nearby, the goal is to move straight to the power up, ignoring any energy blasts. If any close energy blasts close to the player while attempting to retrieve a power up, the player will deploy the shield. The rate of fire will be kept to a minimum for most situations to conserve energy, unless a nearby power up spawns, or if there is only two remaining enemy saucers left. In these cases, the rate of fire will increase, to either deter enemy saucers from retrieving the power up, or attempt to destroy them.

Speed will also be kept at the minimum for most situations, only increasing speed when necessary for dodging, or when a power up spawns nearby, to try and get to it first. Speed will also be increased when there are only two remaining enemy saucers left, attempting to destroy them before the timer runs out.

This controller attempts to implement these strategies with turning, speed, shield deployment and firepower, with the primary goal of flying defensively by dodging energy blasts and retrieving nearby power ups. When there are only two enemies left, the saucer will begin to fly offensively, increasing speed and firepower, and attempt to destroy the enemies before the timer runs out.
\\
\\
The following strategy will be implemented:

\begin{itemize}
\item Fly defensively
	\begin{itemize}
	\item Try to stay away from the ``furball'', let the other saucers destroy each other
	\item Conserve energy
	\item Keep rate of fire to a minimum
	\item Focus movement on dodging energy blasts
	\end{itemize}
\item If a powerup spawns nearby
	\begin{itemize}
	\item Stop dodging energy blasts
	\item Move straight to the powerup at high speed
	\item If there are no close, incoming energy blasts, fire at maximum rate of fire to deter others from reaching the powerup
	\item If there are close, incoming energy blasts, deploy shield
	\end{itemize}
\item If there are only two enemies left
	\begin{itemize}
	\item React to any powerup, even those that are far, to starve the remaining enemies of extra power
	\item Fly aggressively
	\item Increase rate of fire
	\item Attempt to destroy remaining enemies before timer runs out
	\end{itemize}
\end{itemize}